---
title: "Homework 3"
author: "Joe Kingsbury"
output:
  pdf_document: default
  html_notebook: default
---

### **Question 1**
Fit a logistic regression model that assumes the probability of success is an additive function of variables x1 and x2.
```{r}
dat <- read.csv("Homework 3 Data.csv")
fit <- glm(y ~ x1 + x2, data = dat, family = binomial)
summary(fit)
```

### **Question 2**
Interpret the effect of variable x1 on the log odds of success. Verify your interpretation in R.
```{r}
betas <- coef(fit)
p_2 <- plogis(betas[1] + betas[2] * 2 + betas[3] * 0)
p_2

p_3 <- plogis(betas[1] + betas[2] * 3 + betas[3] * 0)
p_3

log((p_3 / (1 - p_3)) / (p_2 / (1 - p_2)))
```

### **Question 3**
Interpret the effect of variable x2 on the log odds of success. Verify your interpretation in R.
```{r}
p_4 <- plogis(betas[1] + betas[3] * 0)
p_4

p_5 <- plogis(betas[1] + betas[3] * 1)
p_5

log((p_5 / (1 - p_5)) / (p_4 / (1 - p_4)))
```

### **Question 4**
Duplicate the Wald Test and p-values for variables x1 and x2 performed by the glm() function. Do you reject or fail to reject your null hypothesis?
```{r}
#Test statistic for x1
x1_ts <- betas[2] / summary(fit)[['coefficients']]['x1', 'Std. Error']
#P-value for x1
2 * pnorm(-1 * abs(x1_ts), mean = 0, sd = 1)

#Test statistic for x2
x2_ts <- betas[3] / summary(fit)[['coefficients']]['x2b', 'Std. Error']
#P-value for x2
2 * pnorm(-1 * abs(x2_ts), mean = 0, sd = 1)

```
For the variable x1 we fail to reject the null hypothesis because the p-value is quite large, > 0.05. However, for the variable x2 we reject the null hypothesis because the p-value is less than our citical alpha level, 0.05 > 0.019.

### **Question 5**
Predict and plot the mean probability of success over the range of values of x1.
```{r}
#Log odds of success
x1 <- seq(from = min(dat$x1), to = max(dat$x1),length.out = 100)
y <- betas[1] + (betas[2] * x1) + (betas[3] * 1)
plot(x = x1, y = plogis(y), ylab = 'Probability of Success',
xlab = 'x1', cex.axis = 1.5, cex.lab = 1.5, type = 'l')
```

