---
title: "Kingsbury Homework 1"
output: html_notebook
---

# WMAN 633 QUANTITATIVE ECOLGOY

## Question 1

### Import the "beach.csv" dataset into R and examine the first few rows of data
```{r}
setwd(choose.dir())
beachdata <- read.csv("beach.csv")
head(beachdata)
```

## Question 2

### Fit a model that includes interactions between Year and BeachID. Report estimates of all coefficients
```{r}
beachfit <- lm(OpenBeach ~ Year * BeachID, data = beachdata)
coef(beachfit)
```

## Question 3

### Predict the beach area at all three beaches in 1992
```{r}
betas <- coef(beachfit)
betas
BeachA <- betas[1] + betas[2] * (1992)
BeachB <- betas[1] + betas[2] * (1992) + betas[3] + betas[5] * 1992
BeachC <- betas[1] + betas[2] * (1992) + betas[4] + betas[6] * 1992
c(BeachA, BeachB, BeachC)
```
#### Beach A = 362.8150, Beach B = 100.0340, Beach C = 214.3744

## Question 4

### Plot the predicted area of each beach across all years
```{r}
ya <- betas[1] + betas[2] * (1992:2014)
yb <- betas[1] + betas[2] * (1992:2014) + betas[3] + betas[5]*(1992:2014)
yc <- betas[1] + betas[2] * (1992:2014) + betas[4] + betas[6]*(1992:2014)

plot(x = beachdata$Year, y = beachdata$OpenBeach, type = 'p', cex.axis = 1.3, cex.lab = 1.3, ylab = "Open Beach Area", xlab = "Year")
lines((1992:2014), ya, col = "blue")
lines((1992:2014), yb, col = "red")
lines((1992:2014), yc, col = "green")
legend("topright", 
  legend = c("Beach A", "Beach B", "Beach C"), 
  col = c("blue", "red", "green"),
  lty = 1)
```

## Question 5

### Interpret the rate of beach area change at each of the 3 beaches



#### Beach A
```{r}
betas
Change_BeachA <- betas[2]
Change_BeachA
```
Beach A is losing 18.36 hectares of open beach per year



#### Beach B
```{r}
Change_BeachB <- betas[2] + betas[5]
Change_BeachB
```
Beach B is losing 1.32 hectares per year.



#### Beach C
```{r}
Change_BeachC <- betas[2] + betas[6]
Change_BeachC
```
Beach C is losing 3.99 hectares per year.
