---
title: "WMAN 633 Homework 7"
author: "Joe Kingsbury"
Date: "April 19th, 2021"
output: html_notebook
---

## **Question 1**
Load data and place into an unmarkedFrameOccu object

```{r, echo=TRUE}
#Set working directory
#setwd(choose.dir())
bobcat <- read.csv("Bobcat.csv") # detection/non-detection data
p_covs <- read.csv("p covariates.csv") #Detection covarites
psi_covs <- read.csv("psi covariates.csv") #Site level covairates

bobcat_mat <- as.matrix(bobcat)

det_covs <- list(
  People = data.frame(p_covs[,c(1:71)]))
head(det_covs$People)
library(unmarked)

occu_data <- unmarkedFrameOccu(y = bobcat_mat, siteCovs = psi_covs, obsCovs = det_covs)
head(occu_data)
```

## **Question 2**
Fit the following candidate set of models:

Detection model |Occupancy model
--------------- | ----------------
intercept-only  |intercept-only
people          |intercept-only
intercept-only  |disturbance
people          |disturbance

```{r, echo=TRUE}
fit1 <- occu(~ 1 ~ 1, data = occu_data)
fit2 <- occu(~ People ~ 1, data = occu_data)
fit3 <- occu(~ 1 ~ Dist_5km, data = occu_data)
fit4 <- occu(~ People ~ Dist_5km, data = occu_data)
summary(fit1)
summary(fit2)
summary(fit3)
summary(fit4)
```

## **Question 3**
Perform model selection with AIC. What is your top model? How do you know? Is there model selection uncertainty?

```{r, echo=TRUE}
library(AICcmodavg)
occu_cand.set <- list(
  F1 = fit1, F2 = fit2, F3 = fit3, F4 = fit4)

occu_mods <- aictab(cand.set = occu_cand.set, second.ord = F)
occu_mods
```
My top model is model 3, fit3, which uses the intercept-only for the detection model and the disturbance within 5km for the abundance model. I know this is my top model because the $\Delta$AIC for Model 1 and Model 2 is greater than 2. There is some uncertainty between model 3 and model because the $\Delta$AIC between Model 3 and Model 4 is less than 2.

## **Question 4**
Average both the effect of people on detection, and disturbance on occupancy, over all models. Report model-averaged slope coefficients and 95% confidence intervals.

```{r, echo=TRUE}
#People's effect on detection across all models
people_avg <- modavgShrink(cand.set = occu_cand.set, parm = 'People', second.ord = F, parm.type = 'detect')

People_slopecoef <- people_avg$Mod.avg.beta
People_slopecoef

people_95 <- cbind(people_avg$Lower.CL,people_avg$Upper.CL)
people_95

#Effect of disturbance on occupancy across all models
dist_avg <- modavgShrink(cand.set = occu_cand.set, parm = 'Dist_5km', second.ord = F, parm.type = 'psi')

Dist_5km_slopecoef <-  dist_avg$Mod.avg.beta
Dist_5km_slopecoef

dist_95 <- cbind(dist_avg$Lower.CL, dist_avg$Upper.CL)
dist_95
```

## **Question 5**
Obtain and plot model-averaged predictions of occupancy probability and detection probability. Average over all models, and make predictions over the observed range of each variable.

```{r, echo=TRUE}
#Detection predictions
p_covs[is.na(p_covs)] = 0 #Necessary because you can't generate new data with NAs
new_p <- data.frame(
  People = seq(min(p_covs[1:71]), to = max(p_covs[1:71]), length.out = 100))

avg_prd_det <- modavgPred(cand.set = occu_cand.set, newdata = new_p, second.ord = F, parm.type = 'detect') 
avg_prd_det

plot(x = new_p$People, y = avg_prd_det$mod.avg.pred, type = 'l', ylab = 'Predicted Detection Probability', xlab = 'People', ylim = c(0, max(avg_prd_det$upper.CL)))
lines(x = new_p$People, y = avg_prd_det$upper.CL, lwd = 1, lty = 2)
lines(x = new_p$People, y = avg_prd_det$lower.CL , lty = 2)

#Occupancy predictions
new_psi <- data.frame(Dist_5km = seq(from = min(psi_covs), to = max(psi_covs), length.out = 100))
new_psi

avg_prd_occupancy <- modavgPred(cand.set = occu_cand.set, newdata = new_psi, second.ord = F, parm.type = 'psi')
avg_prd_occupancy

plot(x = new_psi$Dist_5km, y = avg_prd_occupancy$mod.avg.pred, type = 'l', ylab = 'Predicted Occupancy Probability', xlab = 'Disturbance within 5km')
lines(x = new_psi$Dist_5km, y = avg_prd_occupancy$upper.CL, lwd = 1, lty = 2)
lines(x = new_psi$Dist_5km, y = avg_prd_occupancy$lower.CL , lty = 2)
```

