---
title: "WMAN 633 Homework 5"
author: "Joe Kingsbury"
output:
  pdf_document: default
  html_notebook: default
---

## **Question 1**
Load data and place into an unmarkedFrameOccu object

```{r}
library(unmarked)
bobcat <- read.csv("Bobcat.csv") # detection/non-detection data
p_covs <- read.csv("p covariates.csv") #Detection covarites
site_covs <- read.csv("psi covariates.csv") #Site level covairates

bobcat_mat <- as.matrix(bobcat)

det_covs <- list(
  people = data.frame(p_covs[,c(1:71)])
)


occu_data <- unmarkedFrameOccu(y = bobcat_mat, siteCovs = site_covs, obsCovs = det_covs)
head(occu_data)
```

## **Question 2**
Fit an occupancy model that assumes conditional detection probability is a function of the number of
people photographed at a site, and occupancy probability is a function of disturbance within 5km.

```{r}
fit1 <- occu(~ people ~ Dist_5km, data = occu_data)
summary(fit1)
```

## **Question 3**
Interpret the effect of the number of people photographed at each site on conditional detection probability
of bobat. Verity your interpretation in R.

```{r}
betas <- coef(fit1)
betas

p_1 <- plogis(betas[3] + betas[4])
p_1

p_0 <- plogis(betas[3])
p_0

log((p_1/ (1 - p_1)) / (p_0 / (1 - p_0)))
```
For each person photographed at each site the log-odds of detecting a bobcate decrease by -0.061.

## **Question 4**
Re-create the Wald test for the effect of disturbance within 5km. What is the null hypothesis? Do you
reject or fail to reject this null hypothesis?

```{r}
w <- betas[2] / 4.7728768	
w

p_value <- 2 * pnorm(-1 * abs(w), mean = 0, sd = 1)
p_value
```
The null hypothesis is that disturbance within 5km of our sites has no effect on the conditional probability of detecting at least 1 bobcat. Based on the incredibly small p-value I would go ahead and reject the null hypothesis and disturbance withing 5km of our sampling sites has a significant effect on our detection probability.

## **Question 5**
Predict and plot the effect of disturbance within 5km on bobcat occupancy probability. Do this over
the range of observed disturbance within 5km.

```{r}
new_psi <- data.frame(Dist_5km = seq(from = min(site_covs$Dist_5km),
to = max(site_covs$Dist_5km), length.out = 100))

predict <- predict(object = fit1, newdata = new_psi,
type = 'state')
head(predict)

Dist_5km <- seq(from = min(site_covs$Dist_5km),
to = max(site_covs$Dist_5km), length.out = 100)

plot(Dist_5km, plogis(predict$Predicted), type = "l", ylab = "Predicted Probability of Detection", xlab = "Disturbance within 5km")
lines(x = Dist_5km, y = plogis(predict$lower), lty = 2)
lines(x = Dist_5km, y = plogis(predict$upper), lty = 2)
```

