---
title: "Kingsbury_Homework_4"
author: "Joe Kingsbury"
date: "3/17/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
### **Question 1**
Fit a Poisson regression model that assumes expected count is an interactive function of variables x1 and x2.

```{r}
dat <- read.csv("Homework 4 Data.csv")
head(dat)

fit <- glm(y ~ x1 * x2, family = poisson, data = dat)
summary(fit)
```


### **Question 2**
Interpret the effect of variable x1 on the expected count when x2 is fixed at level "b". Verify your interpretation in R.

```{r}
betas <- coef(fit)
y_1 <- betas[1] + betas[2] * 1 + betas[3] + betas[4] * 1
y_2 <- betas[1] + betas[2] * 2 + betas[3] + betas[4] * 2
exp(y_2) / exp(y_1)
exp(betas[2] + betas[4])
(exp(betas[2] + betas[4]) - 1) * 100
```
The expected count of Y increased by %46.09 for each single unit increase in x1 when taking into account x2 = b.

### **Question 3**
Interpret the effect of variable x2 on the expected count when x1 is fixed at 1. Verify your interpretation in R.

```{r}
y_a <- betas[1] + betas[2] * 1 
y_b <- betas[1] + betas[2] * 1 + betas[3] + betas[4] * 1
log(exp(y_b) / exp(y_a))
betas[3] + betas[4]
(exp(betas[3] + betas[4]) -1) * 100
```
When x1 is held at 1 there is a %43.35 difference between the expected count of y when comparing category a and category b.

### **Question 4**
Predict the expected count, 95% confidence intervals, over the observed range of values of x1, assuming x2 is fixed at level "b".

```{r}
#Create a new dataframe of predicted values
data2 <- data.frame(
x1 = seq(min(dat$x1), max(dat$x1), length.out = 100),
x2 = factor(x = rep('b', times = 100),
levels = c('a','b')))
head(data2)

prd1 <- predict.glm(object = fit, newdata = data2, type = 'link', se.fit = T)
low <- exp(prd1$fit - qnorm(0.975) * prd1$se.fit)
high <- exp(prd1$fit + qnorm(0.975) * prd1$se.fit)

plot(y = exp(prd1$fit), x = data2$x1, xlab = 'x1',
  ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
  ylim = c(min(low), max(high)), type = 'l')
  lines(x = data2$x1, y = low, lty = 2)
  lines(x = data2$x1, y = high, lty = 2)
```


### **Question 5**
Predict the expected count, 95% confidence intervals, of levels "a" and "b", assuming x1 is fixed at it’s mean.

```{r}
#Predicted count of level "a" with x1 fixed at it's mean
y_a_prd <- betas[1] + betas[2] * mean(dat$x1)
exp(y_a_prd)

#Predicted count of level "b" with x1 fixed at it's mean
y_b_prd <- betas[1] + betas[2] * mean(dat$x1) + betas[3] + betas[4] * mean(dat$x1)
exp(y_b_prd)

```
```{r}
#Create a new datafram with x1 at it's mean and level a and b present
data3 <- data.frame(
  x1 = rep(mean(dat$x1), times = 2),
  x2 = c('a', 'b'))

y_prd <- predict.glm(object = fit, newdata = data3, type = 'link', se.fit = T )
y_low <- exp(y_prd$fit - qnorm(0.975) * y_prd$se.fit)
y_high <- exp(y_prd$fit + qnorm(0.975) * y_prd$se.fit)

```


Confidence Interval for "a"
```{r}
Name <- c("Lower Limit", "Upper Limit")
aConfidenceInterval <- cbind.data.frame(y_low[1:1], y_high[1:1])
names(aConfidenceInterval) <- Name
aConfidenceInterval
```

Confidence Interval for "b"
```{r}
bConfidenceInterval <- cbind.data.frame(y_low[2:2], y_high[2:2])
names(bConfidenceInterval) <- Name
bConfidenceInterval
```